### 2021-09-24 README.Rmd
update README with new workflow and figures

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# conmat

<!-- badges: start -->
[![Codecov test coverage](https://codecov.io/gh/njtierney/conmat/branch/master/graph/badge.svg)](https://codecov.io/gh/njtierney/conmat?branch=master)
[![R-CMD-check](https://github.com/njtierney/conmat/workflows/R-CMD-check/badge.svg)](https://github.com/njtierney/conmat/actions)
<!-- badges: end -->

The goal of conmat is to provide methods for producing contact matrices.

## Installation

You can install the development version with:

```r
install.packages("conmat", repos = "https://njtierney.r-universe.dev")
```

Or alternatively you can use `remotes` (although I recommend using the above code)

``` r
# install.packages("remotes")
remotes::install_github("njtierney/conmat")
```

## Example

We can extract out the ABS age population data using `abs_age_lga` like so:

```{r get-syd}
library(conmat)
sydney_age_pop <- abs_age_lga("Sydney (C)")
sydney_age_pop
```

Note that you need to use the exact LGA name - you can look up LGA names in the data set `abs_lga_lookup`:

```{r abs-lga-lookup}
abs_lga_lookup
```

First we want to fit the model to the polymod data

```{r fit-polymod}
set.seed(2021-09-24)
polymod_contact_data <- get_polymod_setting_data()
polymod_survey_data <- get_polymod_population()

setting_models <- fit_setting_contacts(
  contact_data_list = polymod_contact_data,
  population = polymod_survey_data
  )
```

Then we take this model and extrapolate to the sydney data:

```{r sydney-synth-5}
set.seed(2021-09-24)
synthetic_settings_5y_sydney <- predict_setting_contacts(
  population = sydney_age_pop,
  contact_model = setting_models,
  age_breaks = c(seq(0, 85, by = 5), Inf)
  )

synthetic_settings_5y_sydney
```

```{r sydney-synth-5-plot}
# this code is erroring for the moment - something to do with rendering a large plot I think.
# plot_setting_matrices(
#   synthetic_settings_5y_sydney,
#   title = "Setting-specific synthetic contact matrices (Sydney 2020 projected)"
# )
```

```{r plot-matrix-differents}
set.seed(2021-09-24)
plot_matrix(synthetic_settings_5y_sydney$home)
plot_matrix(synthetic_settings_5y_sydney$work)
plot_matrix(synthetic_settings_5y_sydney$school)
plot_matrix(synthetic_settings_5y_sydney$other)
plot_matrix(synthetic_settings_5y_sydney$all)
```

```{r example}
# build synthetic age-structured contact matrices with GAMs
# analysis of polymod data

# set age breaks
age_breaks_5y <- c(seq(0, 75, by = 5), Inf)
age_breaks_1y <- c(seq(0, 100, by = 1), Inf)

# fit a single overall contact model to polymod
m_all <- fit_single_contact_model(
  contact_data = get_polymod_contact_data("all"),
  population = get_polymod_population()
)

m_all
```

```{r syn-all-5y}
# predict contacts at 1y and 5y resolutions for inspection
synthetic_all_5y <- predict_contacts(
  model = m_all, 
  population = get_polymod_population(),
  age_breaks = age_breaks_5y
) %>%
  predictions_to_matrix()

synthetic_all_5y
```

```{r matrix-plot}
set.seed(2021-09-08)
# predict contacts at 1y and 5y resolutions for inspection
synthetic_all_5y <- predict_contacts(
  model = m_all, 
  population = get_polymod_population(),
  age_breaks = age_breaks_5y
) %>%
  predictions_to_matrix()

synthetic_all_5y

library(ggplot2)
plot_matrix(synthetic_all_5y) +
  ggtitle("synthetic all at once")
```

## Speeding up computation with `future`

`conmat` now supports parallelisation, which is useful in a couple of contexts with the model fitting, here is an example:

```{r load-future}
library(future)
plan(multisession, workers = 4)
```

We set the future plan, saying multisession, with 4 workers.

Then we run the code as normal:

```{r show-off-furrr}
polymod_setting_data <- get_polymod_setting_data()
polymod_population <- get_polymod_population()

contact_model <- fit_setting_contacts(
  contact_data_list = polymod_setting_data,
  population = polymod_population
)

contact_model_pred <- predict_setting_contacts(
  population = polymod_population,
  contact_model = contact_model,
  age_breaks = c(seq(0, 75, by = 5), Inf)
)
```

Notably this is about 3 times faster than without using that plan.

## Data sources

This package provides data for use in calculating contact matrices.

These data are still being cleaned and processed, but their current forms are shown below:

### abs_education_state

```{r abs-education-state}
abs_education_state
```

### abs_education_state_2020

```{r abs-education-state-2020}
abs_education_state_2020
```

### abs_employ_age_lga

```{r abs-employ-age-lga}
abs_employ_age_lga
```

### abs_household_lga

```{r abs-household-lga}
abs_household_lga
```

### abs_pop_age_lga_2016

```{r abs-pop-age-lga-2016}
abs_pop_age_lga_2016
```

### abs_pop_age_lga_2020

```{r abs-pop-age-lga-2020}
abs_pop_age_lga_2020
```

### abs_state_age

```{r abs-state-age}
abs_state_age
```

## Code of Conduct

Please note that the conmat project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.