### 2024-08-08 paper/paper.Rmd
updates from meeting with Mike

---
title: 'conmat: generate synthetic contact matrices for a given age-stratified population'
authors:
- affiliation: 1
  name: Nicholas Tierney
  orcid: 0000-0003-1460-8722
- affiliation: 1,2
  name: Nick Golding
  orcid: 0000-0001-8916-5570
- affiliation: 1,3
  name: Aarathy Babu
  orcid: 
- affiliation: 4
  name: Michael Lydeamore
  orcid: 0000-0001-6515-827X
- affiliation: 1,3
  name: Chitra Saraswati
  orcid: 0000-0002-8159-0414
date: "03 May 2024"
output:
  html_document:
    keep_md: yes
  pdf_document: default
bibliography: references.bib
tags:
- epidemiology
- R
- infectious disease
affiliations:
- index: 1
  name: Telethon Kids Institute
- index: 2
  name: Curtin University
- index: 3
  name: 
- index: 4
  name: Monash University
---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
#| comment: false
knitr::opts_chunk$set(comment = "#>",
                      echo = TRUE,
                      out.width = "95%",
                      retina = 3,
                      fig.height = 4,
                      fig.align = "center",
                      dpi = 300,
                      dev = "png")
options(tinytex.clean = FALSE)
```

# Summary

[ A summary describing the high-level functionality and purpose of the software for a diverse, non-specialist audience. ]

This article introduces `conmat`, an R package which generates synthetic contact matrices. Contact matrices describe the number of contacts between individuals, typically age groups. They are used to create models of infectious disease spread.

There are currently only a few options for a user to access synthetic contact matrices [@socialmixr; @prem]. Existing methods describing how to generate synthetic contact matrices from @prem are not designed for replicability, and not all areas of interest (countries) are included.

[ What's different and useful about conmat? ]

Current interfaces to contact matrices are typically fixed to a given country. There is not a designed interface for someone to access a contact matrix for a sub-region of a country, such as a state, city, or suburb. These sub-region areas are important, as many public health decisions are made at that level.

The conmat software allows users to generate their own contact matrix, by providing their own age population data and combining this with information from an existing contact survey, such as POLYMOD [@polymod]. The software exposes model fitting and prediction separately to the user. Users can fit a model based on a contact survey such as POLYMOD, then predict from this model to their own data. This means users can generate synthetic contact matrices for any region, with any contact survey, provided they have the relevant data.

[ What else is covered in this paper? ]

We demonstrate a use-case for `conmat` by creating contact matrices for a state (i.e. at the sub-national level) in Australia.

[Need to work out where to fit in `syncomat`]

It is also worth noting that there is an updated analysis pipeline that uses conmat, [`syncomat`](https://github.com/idem-lab/syncomat), to generate synthetic contact matrices for 200 countries.

# Statement of need

[ A Statement of need section that clearly illustrates the research purpose of the software and places it in the context of related work. ]

#TODO - A better first sentence that encapsulates conmat use?

Infectious diseases like influenza and COVID19 spread via social contact. If we can understand patterns of contact - which people are more likely be in contact with each other - then we will be able to create models of how disease spreads. Epidemiologists and public policy makers can use these models to make decisions to keep a population safe and healthy.

Identifying vulnerable groups and predicting how diseases spread are essential for informed public health decision-making. Infectious diseases such as influenza and coronavirus spread through human-to-human interactions, or in other words, "social contact". Quantifying social contact and its patterns can provide critical insights into how these diseases spread. [ Is this circular? ] / and how best to mitigate the spread of these diseases.

Social contact surveys quantify social contact by asking people to describe the frequency and type of social contact across different settings (home, work, school, other). These surveys provide (?) a measure of contact rates: an empirical estimate of the number of social contacts from one age group to another and the setting of contact. For example, we might learn from a contact survey that homes have higher contact between 25-50 year olds and 0-15 year olds, whereas workplaces might have high contact within 25-60 year olds.

These social contact surveys exist for a few countries. As an example, the "POLYMOD" study by @mossong2008 covered 8 European countries: Belgium, Germany, Finland, Great Britain, Italy, Luxembourg, The Netherlands, and Poland [@mossong2008]. However, what do we do when we want to estimate contact rates in other countries where this is not yet measured? We can use existing data--the contact rates obtained from contact surveys--to help us predict these estimates to countries or places that do not have them. These are called "synthetic contact matrices". A popular approach by @prem2017 predicted contact rates from the POLYMOD study to 152 countries. This was later updated to include synthetic contact matrices for 177 countries at "urban" and "rural" levels for each country [@prem2021].

However, there were major limitations with the methods in @prem2021. First, not all countries were included in their analyses. Second, some of the synthetic contact matrices did not have enough granularity; in other words, they covered areas that are too large, such as the "urban" or "rural" parts of a country. This is disadvantageous as public health groups might need to make predictions for more fine-grained areas within a country, such as a district or municipality. Third, the methodology used by Prem et al. was challenging to reuse in other contexts. Prem et al. provided the code used for their analysis, but that code was not designed for replicability and easy modification with user-defined inputs.

[REVISED PARAGRAPH BELOW] The `conmat` package was created to fill a specific need for creating synthetic contact matrices for specific local government areas for Australia, for work commissioned by the Australian government. We created methods and software to facilitate the following:

The `conmat` package was developed to fill the specific need of creating synthetic contact matrices for local government areas in Australia. This package is used for [ #TODO what work, specifically? Health? Provide example. Or is *this* package commissioned by the Aus govt? note: @MikeLydeamore will reference the official government report] work commissioned by the Australian government.
We developed methods and software to facilitate the following tasks.

- Generate, as output, synthetic contact matrices from age-stratfied population data.
- Create next generation matrices (NGMs).
- Apply vaccination reduction to NGMs.
- Use NGMs in disease modelling.
- Provide tidied population data from the Australian Bureau of Statistics.

# Example

As an example, let us generate a contact matrix for a local government area within Australia, using a model fitted from the POLYMOD data.

Suppose we want to generate a contact matrix for the City of Perth. We can get the age-stratified population data for Perth from the helper function `abs_age_lga`:

```{r}
#| label: load-conmat
library(conmat)
perth <- abs_age_lga("Perth (C)")
perth
```

We can then generate a contact matrix for `perth` using the `extrapolate_polymod` function, where the contact matrix is generated using a model fitted from the POLYMOD data.

```{r}
#| label: extrapolate-polymod
#| echo: true
perth_contact <- extrapolate_polymod(population = perth)
perth_contact
```

We can plot the resulting contact matrix for Perth with `autoplot`:

```{r}
#| label: autoplot-contacts
autoplot(perth_contact)
```

# Implementation

`conmat` was built to predict at four settings: work, school, home, and other. 
One model is fitted for each setting. 
Each model fitted is a Poisson generalised additive model (GAM) which predicts the count of contacts, with an offset for the log of participants. 
The model has six covariates to explain six key features of the relationship between ages, 
and two optional covariates for attendance at school or work.
The two optional covariates are included depending on which setting the model is fitted for.

Each cell in the resulting contact matrix, indexed *i*, *j*, is the predicted number of people in age group *j* that a single individual in age group *i* will have contact with per day. If you sum across all the *j* age groups for each *i* age group, you get the predicted total number of contacts per day for each individual of age group *i*. [ #TODO expected, predicted, or average? Does it matter? note: currently going with predicted]

The six covariates are 
$|i-j|$, 
${|i-j|}^{2}$, 
$i + j$, 
$i \times j$, 
$\text{max}(i, j)$ and 
$\text{min}(i, j)$.

The six key features of the relationship between the age groups, represented by the six covariates, are displayed in the figure below. 
[ #TODO notes-to-self: the model structure wasn't generated through any particularly robust process, it was just coming up with structures that looked mildly appropriate for our use case. ]

```{r}
# Show partial dep plot of the six main covariates
```

Note that these partial dependency plots are on the log scale. 
When the six covariates are added up together for each setting (in other words, each model) and exponentiated, they show the following patterns:

```{r}
# Show combined partial dep plot (i.e. sum of the partial dependencies for all six covariates) in each setting: home, school, work and other
```

In other words, the six covariates above provide patterns that are useful in modelling the different settings,
and correspond with real-life situations of how contact would look like. 
In the home setting for example, [ #TODO describe how children interact with parents and elderly generation, grandparents ].
When the covariates for school and work are added, these covariates also provide patterns that correspond with real-life situations.
https://idem-lab.github.io/conmat/dev/articles/visualising-conmat.html
In the school setting, children tend to contact other children in the same age groups as them.
In the work setting, there are no contacts with children under the age of ten and minimal contact with adults beyond retirement age.

One of the issues with the contact matrices generated by @prem2017 is that some countries are missing. To remedy this we generated synthetic contact matrices for 200 countries, based on a list of country names by the UN, fitted on the POLYMOD contact surveys.
We also ensured that the analysis pipeline is reproducible and transparent by utilising a targets workflow, which allows ease of editing for users.
The resulting synthetic contact matrices, and a replicable / extensible (?) analysis pipeline, can be found in the syncomat analysis pipeline ([GitHub](https://github.com/idem-lab/syncomat), [Zenodo](https://zenodo.org/records/11365943)).

## Model interfaces

We provide functions for model fitting at various use cases. Further detail for each of the following functions are available at: https://idem-lab.github.io/conmat/dev/

* `fit_single_contact_model()`
    * Fits a generalised additive model (GAM) using contact survey data and population size information. This function is recommended when you want to fit a model to only one setting, for which you might want to provide your own contact survey data.

* `predict_contacts()`

    * This takes a fitted model from `fit_single_contact_model()` and predicts [ #TODO what is predicted? ] to a provided (?) population structure.

* `fit_setting_contacts()`
    * Fits the `fit_single_contact_model()` to each setting. This function is useful for when you have multiple settings to fit. Returns a list of fitted models.

* `predict_setting_contacts()`
    * Takes a list of fitted models from `fit_setting_contacts()` and predicts [ #TODO what is predicted? ] to a given population for each setting.

* `estimate_setting_contacts()`
    * A convenience function that fits multiple models, one for each setting. This means fitting `fit_setting_contacts()` and then `predict_setting_contacts()`. Recommended for when you have multiple settings to fit and want to predict to a given population as well.

* `extrapolate_polymod()`
    * Takes population information and predicts pre-fit model from POLYMOD - used for speed when you know you want to take an already fit model from POLYMOD and just fit it to your provided population.

[ #TODO for the above it's good to explain what exactly is predicted. Otherwise it's confusing for the user to understand what each of the model outputs? ]

# Conclusions and future directions

The `conmat` software provides a flexible interface to generating synthetic contact matrices using population data and contact surveys. These contact matrices can then be used in infectious disease modelling and surveillance.

## Strengths + Weaknesses

The main strength of `conmat` is its interface requiring only age population data to create a synthetic contact matrix. Current approaches provide only a selection of country level contact matrices. This software can predict to arbitrary demography, such as sub-national, or simulated populations.

We provide a trained model of contact rate that is fit to the POLYMOD survey for ease of use. The software also has an interface to use other contact surveys, such as @comix. This is important as POLYMOD represents contact patterns in 8 countries in Europe, and contact patterns are known to differ across nations and cultures.

The model structure (covariates?) provided in conmat was designed to represent the key features that are typically present in a contact matrix for different settings (work, school, home, other). Including other sources of information that may better describe these contact patterns, such as ..., may improve model performance.

The interface to the model formula in conmat is fixed; users cannot change the covariates of the model. This means that if there is an unusual structure in their contact data it might not be accurately captured by conmat. It was a design decision that was made to focus on the key feature of conmat: using just age population data to predict a contact matrix.

## Meaning of the work

Public health decisions are often based on age specific information, which means the more accurate your age specific models are, the better those decisions are likely to be. This is the first piece of software that will provide appropriate contact matrices for a population, which means more accurate models of disease.

This work was used as a key input into several models for COVID19 in Australia and contributed directly to decisions around vaccination policy.

## Unanswered questions / Future Directions

Future directions for this software could include:

* Create a contact matrix using a custom contact survey from another source, such as  the `socialmixr` R package.
* Predict to any age brackets - such as monthly ages, for example, 1, 3, 6, month year old infants
* Add ability to fit multiple contact surveys at once, e.g., POLYMOD and another dataset
* Add ability to include known household age distributions as offsets in the 'home' setting model, in place of the whole population distribution. So compute household age matrices (like age-structured contact matrices, but for household members instead of contacts) from POLYMOD data. If we compute a different one for each household size, in the POLYMOD data (probably estimated with another GAM, to make best use of the limited data) we might be able to extrapolate household age matrices to new countries based on the distribution of household sizes.
* Add methods for including household size distributions
* Add uncertainty to estimates
* Move Australian centric data into its own package
* Add documentation on specifying your own GAM model and using this workflow

Software is never finished, and the software in its current format has proven useful for infectious disease modelling. In time we hope it can become more widely used and be useful for more applications in epidemiology and public health.

# References

https://www.doherty.edu.au/our-work/institute-themes/viral-infectious-diseases/covid-19/covid-19-modelling/modelling

# Cutting Room Floor

`conmat` also provides flexibility, in that it allows users to specify the area in which they would like the contact matrices to be generated; it allows users to specify their own age groups and population structures; it allows users to upload their own contact surveys to fit the model on; and it allows users to generate the contact matrices at different settings.